RationalGRL: A Framework for Argumentation and Goal Modeling

### 2018-05-04 Response to reviewers.txt

######## EDITOR ########

===== COMMENT #1
Please ensure also that your manuscript shows clearly how your work relates to relevant works as they may have been published in the Requirements Engineering or presented at the Requirements Engineering series of conferences.

===== REPLY #1
Currently 6 of the 44 references in the article come from either REJ or the RE conference [11,24,26,32,43,47]. We have searched through the journal and conference database for more references, and we found 4 more articles that are related to our work [20,22,27,39]. We have added a discussion of these articles to the related work section, at the beginning of the paragraph "Requirements Engineering". This means we currently have 10 references from the Requirements Engineering Journal or presented at the Requirements Engineering series of conferences.

@@

===== COMMENT #2
The proposed algorithms are very simple and they do not add much to the overall presentation, I think they can be summarized in the section describing the tool.

===== REPLY #2
We agree that many algorithms are similar, and we have therefore shortened the sections containing these algorithms: sections 5.4.1 to 5.4.3 now all contain one example of the specific type of algorithm instead of two or more.

===== COMMENT #3
The methodology and the tools looks quite preliminary and without a real validation of the approach. I think this is the weakest part of the paper. Authors should demonstrate the applicability, scalability and usefulness of their approach in the real word. The tool and the modeling language should be used and evaluated by practitioners while they cope with real requirements engineering problems and complexity issues.

===== REPLY #3
We agree with the reviewer that a real validation of the approach is a necessary next step. However, we would like to point the reviewer to the case study in section 3, in which 153 pages of transcripts of requirements discussions were used (see also the original materials at www.rationalgrl.com). We do agree that further validation of the tool and method is necessary, preferably in an industrial setting. However, such an industrial strength case study is a significant amount of work (see for example Estrada et al. (2006 "An empirical evaluation of the i* framework in a model-based software generation environment". They took 9 months to conduct the case study.), and would most likely yield enough new insights for a separate paper.

@@

===== REPLY #2
We have included in the abstract and introduction that we use a simplified version. Note that all the core elements of GRL are included but, for example, we have included only one level of contribution strength, where GRL contains different levels of positive and negative contribution. This simplification does not in any way influence the case study or any of the other results. Furthermore, extending the GRL part of RationalGRL to include the full set of GRL elements would not change the essence of our formalism.

===== COMMENT #3
I think that the paper should compare the proposed argumentation structure at least with the beliefs in GRL. In general, how does this argumentation approach really differ from one based on beliefs and, in particular, belief revision? Assume that an argument is a belief of some stakeholder revised by another argument, which is a belief of some other stakeholder?

===== REPLY #3
We have added a comparison between arguments in RationalGRL and beliefs in GRL in the related work section, at the beginning of the paragraph "Goal Modelling".

The connection between argumentation and belief revision is interesting, especially from a theoretical point of view, and in knowledge representation and reasoning this is a popular topic of research (see references 5-7 below). However, since studying relations between argumentation and belief revision does not seem directly related to our current framework and seems more interesting from a formal logical point of view, we have left this out of the discussion in our paper, instead choosing to focus on beliefs in GRL.

===== COMMENT #4 & #5
How would you compare this specific approach with the old gIBIS?
Please discuss why you are trying to have detailed, while simple, rationale, although the recent trend in DR is otherwise?

===== REPLY #4 & #5
We have added some discussion in the 2nd paragraph of section 8.1-Design Rationale. A brief comparison with, for example gIBIS, is already in section 8.1-Design Rationale (reference [38] includes gIBIS). We felt that an extensive comparison between the ideas behind goal modelling, which we extend with RationalGRL, and design rationale diagramming methods like gIBIS is not within the extent of this paper, where the focus is more on the connection between "normal" goal modelling (GRL) and extended "argumentative" goal modelling (RationalGRL).

@@

===== COMMENT #6
I wonder about the formalization in propositional logic. Why is it not used for automated reasoning (at least partly instead of the specific algorithms)?

===== REPLY #6
Propositional logic in itself does not contain specific algorithms for, for example, INTRO, REPLACE and DISABLE. Furthermore, the algorithms also introduce arguments and attacks, formal notions of which are not in basic propositional logic. The idea is that the formalization is needed for the algorithms, but that just providing a formal semantics of GRL does not directly tell you how to compute/reason with GRL models.

===== COMMENT #7
I am also missing a discussion about the consequences of this binary "nature" of argumentation with respect to argumentation in reality.

===== REPLY #7
While there is extensive work on less "binary" frameworks for argumentation (see e.g below references 2 and 3), we feel that a discussion on this would not be suitable for the current paper (as it is more of an AI subject than relevant for RE). For an extensive discussion on the relation between the formal argumentation proposed in our article and argumentation in reality, we refer the reviewer to reference [4] below. Note that we have opted not to include these references in the paper, as we are already at 50 references and want to focus more on an RE/SE audience. If the reviewer desires, however, we can add the references to the article.

@@

===== REPLY #11
We agree with the reviewer that the two mentioned sentences are confusing. We have improved this as follows:
- We have rewritten the first sentence of the abstract as follows: "Goal-oriented requirements modeling approaches aim at capturing intentions of various stakeholders involved in the development of an information system and refining them into low level goals and tasks."
- We have deleted the first sentence of the introduction, since we feel that it is not necessary to explain the definition of Requirements Engineering in the Requirements Engineering Journal.

===== COMMENT #12
Finally, you should run a spell checker, have a careful proof reading of grammar, and make sure that there is no unresolved reference to a figure in a submitted manuscript.

===== REPLY #11
We’ve ran a spell-checker and proof-read the paper carefully. There was one unresolved reference to a figure, which a corrected. Thank you for pointing this out.

######## REFERENCES ########

[1] Estrada, H., Rebollar, A. M., Pastor, O., & Mylopoulos, J. (2006, June). An empirical evaluation of the i* framework in a model-based software generation environment. In International Conference on Advanced Information Systems Engineering (pp. 513-527). Springer, Berlin, Heidelberg.
[2] A. Hunter and M. Thimm (2017) Probabilistic Reasoning with Abstract Argumentation Frameworks, Journal of Artificial Intelligence Research, 59:565--611. 
[3] Grossi, D., & Modgil, S. (2015, July). On the Graded Acceptability of Arguments. In IJCAI (pp. 868-874).
[4] van Eemeren, F.H., & Verheij, B. (2017). Argumentation Theory in Formal and Computational Perspective. IFCoLog Journal of Logics and Their Applications 4 (8), 2099-2181.

[5] Falappa, Marcelo Alejandro, Gabriele Kern-Isberner, and Guillermo Ricardo Simari. "Belief revision and argumentation theory." Argumentation in artificial intelligence. Springer, Boston, MA, 2009. 341-360.
[6] Dix, Jürgen, et al. "Belief change and argumentation in multi-agent scenarios (Dagstuhl seminar 13231)." Dagstuhl Reports. Vol. 3. No. 6. Schloss Dagstuhl-Leibniz-Zentrum fuer Informatik, 2013.
[7] Paglieri, Fabio, and Cristiano Castelfranchi. "Revising beliefs through arguments: Bridging the gap between argumentation and belief revision in MAS." International Workshop on Argumentation in Multi-Agent Systems. Springer, Berlin, Heidelberg, 2004.

### 2018-05-04 discussion.tex
final comments FB fixed

Software design and requirements engineering are very closely related \cite{nuseibeh2001weaving} and hence the insights from the DR literature are directly applicable to RE. The RationalGRL framework essentially incorporates the core ideas from DR into goal-oriented requirements engineering by explicitly including arguments for and against the various options into the goal model, and by proposing a development process and critical questions that encourage reflection when thinking about the possible goals and functionality of a system.

>>>>

Software design and requirements engineering are very closely related \cite{nuseibeh2001weaving} and hence the insights from the DR literature are directly applicable to RE. The RationalGRL framework essentially incorporates the core ideas from DR into goal-oriented requirements engineering by explicitly including arguments for and against the various options into the goal model. Note that we explicitly connect to the extensive literature on goal modeling, and in this respect may seem to go against the trend in DR research to diagram decisions and processes. However, we note that RationalGRL also includes a development process and critical questions that encourage reflection when thinking about the possible goals and functionality of a system, which connects with the more research ideas on DR \cite{razavian2016two,SchriekEtal2016}.

### 2018-05-04 validation.tex
final comments FB fixed

\subsection{Evaluation with Enterprise Architects}

\todo{marcvanzee}{marcvanzee,floris,sepideh}{This is copied from the CAiSE paper where we did the original evaluation with two Enterprise Architects. I'm not sure if we can re-use it. The main problem is that at that time we developed GRL models specifically for the organization in which the Enterprise Architects were working, and we don't have such a model now. What do you think?}

To ensure GRL models captured the essence of the EA principles correctly, we evaluated the models with two Enterprise Architects from eGovernment. The
architects found the GRL models very useful in that they could visualize the principles and were easier for them to understand the links between the rationale, added value, actions and the goals of the principles. Furthermore, since the EA
principles were linked to the high-level goals of the organization, it made them easier to justify and analyze. Also, we captured the traceability links between the related EA principles which helped the architects to document the links and
verify the connections.

The architects also informed us that assigning values for contribution links can be done at design time together with the architects.
We aim to extend our models to more principles and provide analysis of the principles with high-level goals as well as low-level operationalized activities of the organization.

>>>>

%FB: I'd leave this out

%\subsection{Evaluation with Enterprise Architects}
%
%\todo{marcvanzee}{marcvanzee,floris,sepideh}{This is copied from the CAiSE paper where we did the original evaluation with two Enterprise Architects. I'm not sure if we can re-use it. The main problem is that at that time we developed GRL models specifically for the organization in which the Enterprise Architects were working, and we don't have such a model now. What do you think?}
%
%To ensure GRL models captured the essence of the EA principles correctly, we evaluated the models with two Enterprise Architects from eGovernment. The architects found the GRL models very useful in that they could visualize the principles and were easier for them to understand the links between the rationale, added value, actions and the goals of the principles. Furthermore, since the EA principles were linked to the high-level goals of the organization, it made them easier to justify and analyze. Also, we captured the traceability links between the related EA principles which helped the architects to document the links and verify the connections.

%The architects also informed us that assigning values for contribution links can be done at design time together with the architects. We aim to extend our models to more principles and provide analysis of the principles with high-level goals as well as low-level operationalized activities of the organization.
